---
title: "zillow cleaning"
author: "Sarah Nash"
date: "`r Sys.Date()`"
output: html_document
---
## Cleaning minumum wage data 
```{r}
wages <- read.csv("Minimum Wage Data.csv") 
w <- wages %>%
  select("Year", "State", "FedMW" = "Federal.Minimum.Wage", "StateMW" = "Effective.Minimum.Wage") %>%
  mutate(StateAbbr =  state.abb[match(State,state.name)])


```
 Not much to clean; write to csv
```{r}
write.csv(w, "MinWage_clean.csv")
```

## Cleaning 2-bedroom zillow house price data 
```{r}
df <- read.csv("zillow_state_sales_2bed.csv")

dates <- colnames(df)[6:ncol(df)]

# Change the columns names to be iso format.
colnames(df)[6:ncol(df)] <- lapply(dates, function(x){str_replace_all(substr(x, 2,11), "\\.", "-") })
```

Convert to long format 
```{r}
df_long <- df %>%
  gather(date, price, `2000-01-31` : `2023-04-30`) %>%
  select(-c(RegionID, SizeRank, RegionType, StateName)) %>%
  rename(State=RegionName)

  
df_long  
```

Write to a csv 
```{r}
# write.csv(df_long, "zillow_sales_2bed_clean.csv")
```


## Cleaning the single family home data 

Import the dataset 
```{r}
df2 <- read.csv("zillow_state_sales.csv")
```

Change the date format to ISO 
```{r}
dates <- colnames(df2)[6:ncol(df2)]
colnames(df2)[6:ncol(df2)] <- lapply(dates, function(x){str_replace_all(substr(x, 2,11), "\\.", "-") })
```

Convert to long format 

```{r}
df2_long <- df2 %>%
  gather(date, price, `2000-01-31` : `2023-04-30`) %>%
  select(-c(RegionID, SizeRank, RegionType, StateName)) %>%
  rename(State=RegionName)

  
df2_long  
```

Write to a new csv 
```{r}
write.csv(df2_long, "zillow_sales_clean.csv")
```

##

Worth of income in relation to housing prices 
Metrics: Hours on given wage to pay off house (bad) 
Number of years at a wage to pay off a house? 
Percent of wage to pay off house (0% interest) in 30 years.
Number of years to buy a house at 33% of wage? 
% of income to pay in 30 years. 
```{r}
test <- df2_long[df2_long$State=="Florida",] %>%
  filter(grepl("-01-31", date)) %>%
  mutate("date" = substr(date, 1, 4)) %>%
  rename("Year" = "date")
  # mutate("Year" = substr(date, 1, 4))
```


```{r}

wage_df <- read.csv("MinWage_clean.csv")
w <- wage_df %>% 
  filter (State == "New York") %>%
  select("Year","StateMW") # %>%
  # plot_ly(type = "scatter", mode = "lines", x =~Year, y =~ State.Minimum.Wage*52*40)
```

```{r}
combined <- merge(test, w, by="Year") 
```

```{r}
plot_ly(
  data = combined, 
  type = "scatter", 
  mode = "line",
  x = ~Year, 
  y = ~price/(StateMW*52*40*30) # This is prop. of income for a 30 year, 0 interest mortgage. 
)
```




## Cleaning the median income data
```{r}
incomedf <- readxl::read_xlsx("income_by_state.xlsx", skip=7) 
incomedf2_current <- incomedf[c(2:53),] %>% select(1, seq(2,ncol(incomedf), by=2)) 
incomedf2_2020 <- incomedf[c(57:nrow(incomedf)),] %>% select(1, seq(2,ncol(incomedf), by=2)) 
```
Convert to a long format of (State, Year, \$ Current, \$ 2020 ), 
also need to chop off col numbers from the year column.... 
```{r}
inc_cur_long <- gather(incomedf2_current, Year, 'Median Income(Current)', "2020":"1984 (19)")
inc_2020_long <- gather(incomedf2_2020, Year, 'Median Income(2020)', "2020":"1984 (19)")

median_income_df <- merge(inc_cur_long, inc_2020_long, by=c("State", "Year"))
median_income_df$Year <- str_split_i(median_income_df$Year, " ",1)
```
Restrict to information I need for the shiny app
```{r}
mi_df <- median_income_df %>%
  filter(Year %in% c(2000:2023)) %>%
  select(-c("Median Income(2020)")) %>%
  rename("Med.Income" = "Median Income(Current)") %>%
  mutate(StateAbbr =  state.abb[match(State,state.name)])

```
Save to a csv 
```{r}
write.csv(mi_df, "MedIncome_clean.csv")
```


## Test Plots 
```{r}
state <- c("Florida")
# ifelse(input$`2brselectpage4` == "family", mydata <- housingdf, mydata <- housingdf.2br)

plotdata <-  df_long %>% filter(State %in% state) %>% na.omit()
plot(x=time(plotdata$date), y=plotdata$price)
abline(reg=lm(price ~ time(date), data=plotdata))

plotdata$lm <- plotdata %>%
  lm(price ~ time(date),.) %>%
  fitted.values()


colors <- c("red", "blue", "green", "pink", "black")
colors <- colors[1:length(state)]
m <- lm(price~date,data = plotdata)
p <- plot_ly(data = plotdata,
             type = "scatter", 
             mode = "lines", 
             color = ~State,
             colors = colors,
             # trendline = "ols",
             x = ~date, 
             y = ~price)
p %>% add_trace(x =~date, y=~lm, mode="lines")
```

## Another Test Plot 
```{r}
test <- df2_long %>%
  filter(grepl("-01-31", date)) %>%
  mutate("date" = substr(date, 1, 4)) %>%
  rename("Year" = "date")
scatter_df <- merge(test, median_income_df, by=c("Year", "State")) %>% na.omit()
cust<- scatter_df[scatter_df$Year==2010,]$price / as.double(scatter_df[scatter_df$Year==2010,]$`Median Income(Current)`)
plot_ly(
  data = scatter_df[scatter_df$Year == 2010,], 
  x = ~price, 
  y = ~`Median Income(Current)`,
  type = "scatter",
  mode = "markers",
  color = ~ cust,

  marker = list(
    # colorscale="Reds",
    size = 12,
    line = list(color = "black", width = 0.5),
    reversescale =TRUE,
    showScale = TRUE
    ),
    colors="YlOrRd"


    # opacity = ~ price / as.double(`Median Income(Current)`)
  # opacity=0.5,
  # colors = colorRamp(c("lightblue", "pink"))

) %>%
  colorbar(title='Affordability',orientation='h',
           tickmode="array", ticktext = c("low", "medium", "high"),
           tickvals=c(min(cust), (min(cust) + max(cust))/2, max(cust)),
           x=0.4, y= -0.3) %>%
  layout(xaxis= list(title="House Price",tickprefix = '$'),
         yaxis = list(title="Median Income", tickprefix='$', tickformat='$' ))

# ) %>% hide_colorbar()

```


```{r}
library(bannerCommenter)
banner("Data Input" , emph = FALSE,bandChar = "-")
open_box("test")
```

